Versioned data: why it’s needed and how it can be achieved (easily and cheaply)

### 2017-08-03 ms.tex
more writing

@@

Tools for creating and archiving digital objects have seen rapid progress in recent years.  At present, each version of the dataset can be automatically archived and assigned a DOI, via several providers (Table \ref{tab:doi_minting}).  Here we describe the particular integration of github and zenodo.  It is useful to have the ability to either be specific---reference a parti

TODO: FLESH THIS OUT AND ADD REFERENCE TO BLOG POST AND OTHER THINKING ON THIS

@@

At current there are two classes of data users: those that  interact with the data via point and click downloading and those that use code-based interaction.  The system we describe allows for both types (Figure \ref{fig:technology_stack} and Table \ref{tab:user_requirements}).  GitHub has a feature called ``releases'' in which specific commits can be turned into a version for distribution (with associated version number).  Those releases can be downloaded directly by users via point-and-click, or accessed via the Github API. Releases can also be embedded into any kind of website that allows the data curator to facilitate discovery by more users and to provide context.  Releases can also be linked to code-based users via an R package which we describe in the following section.

@@

A point-and-click workflow may create barriers for reproducibility \cite{Wilson-2014,Lowndes-2017}. Ideally, users should be able to also able access the all versions of the database programmatically. To this end, we offer a novel implementation of a VDD system. We have written a R package, \texttt{datastorr} (\smurl{github.com/richfitz/datastorr}), that a) interacts with the GitHub API to pull down versions of the dataset; and b) constructs the shell of a second, database specific R package, which can be used to access versions of the dataset. Using this system, a researcher can create and distribute a R package that facilitates access to her data with (very) minimal computational skills.

@@

>>>>

TODO turn this into a paragraph about collaboration

@@

@@

At current there are two classes of data users: those that  interact with the data via point and click downloading and those that use code-based interaction.  The system we describe allows for both types (Figure \ref{fig:technology_stack} and Table \ref{tab:user_requirements}).  GitHub has a feature called ``releases'' in which specific commits can be turned into a version for distribution (with associated version number).  Those releases can be downloaded directly by users via point-and-click, or accessed via the Github API. Releases can also be embedded into any kind of website that allows the data curator to facilitate discovery by more users and to provide context.  Releases can also be linked to digital object identifiers and code-based users via for example an R package.

@@

A point-and-click workflow may create barriers for reproducibility \cite{Wilson-2014,Lowndes-2017}. Ideally, users should be able to also able access the all versions of the database programmatically. Code to access the Github releases could be written individually by each user, but this creates an unnecessary technological hurdle for reproducibility.  To make access via code as easy as possible for users, we offer a novel implementation of a VDD system. We have written a R package, \texttt{datastorr} (\smurl{github.com/richfitz/datastorr}), that a) interacts with the GitHub API to pull down versions of the dataset; and b) constructs the shell of a second, database specific R package, which can be used to access versions of the dataset. Using this system, a researcher can create and distribute a R package that facilitates access to her data with (very) minimal computational skills.

@@

Systematically archiving data is a key part of good workflow \citep{Wilkinson-2016, Piwowar-2011, Whitlock-2011}, and tools for creating and archiving digital objects have seen rapid progress in recent years.  At present, each version of the dataset can be automatically archived and assigned a DOI, via several providers (Table \ref{tab:doi_minting}).  Here we describe the particular integration of github and zenodo.  It is useful to have the ability to either be specific---reference a particular version for the purposes of reproducibility or general--reference the entire project.  Zenodo currently automatically supplies both ``concept'' and ``version'' DOI's to the commits associated with Github releases automatically \citep[Figure \ref{fig:semantic} and ][]{Nielsen-2017}.  This allows users to very precisely reference either generally or specifically, both of which may be appropriate in different situations.  There is an unresolved issue about properly allocating credit to authors for these citations, and we hope that the large citation mapping services (Web of Science and Google Scholar) will adapt to the community standard once the research community settles on best archiving practices for versioned data.
